Response to the defense feedback

### 2021-06-19 Chapter-2/Chapter-2.tex
Updates from Overleaf

According to Table~\ref{table:side_channel_detection_tool}, some methods may choose to analyze the source code or the intermediate representations of the vulnerable software. However, it still cannot solve the problem. First, many widely-used libraries (e.g., Intel MKL) do not provide source code. Second, some libraries use hard-coded assembly code to pursue maximum performance.
CacheAudit~\cite{182946} uses abstract domains to compute an
over-approximation of cache-based side-channel information leakage upper bound.
However, it is difficult to judge the sensitive level of the side-channel leakage based on the leakage provided by CacheAudit. Besides, CacheAudit is not scalable enough to analyze ciphers like RSA implementations. CacheS~\cite{236338} improves the approach from
CacheAudit with a new abstract model that only tracks
secret-related code. Like CacheAudit, CacheS cannot
measure the amount of leaked information in each side-channel vulnerability.
CaSym~\cite{Brotzman19Casym} introduces a static cache-aware symbolic reasoning technique to cover multiple paths in a target program. CaSym can work on two threat models: trace-based attacks and access-based attacks. Again, their
approaches cannot evaluate the sensitive level for each side-channel vulnerability, and it only works on small code snippets.

@@

\subsubsection*{Software Patch}
Once we have identified side-channel leakages, there are two common approaches to mitigate side-channel attacks. The first approach is to adopt unified behavior code when computing the secret data. For example, a control flow leakage happens when different secret inputs lead to different paths. Software developers can mitigate the leakage by executing both branches and select the result with bit-wise operations~\cite{Coppens:2009:PMT:1607723.1608124}. However, for exiting code base, it is hard to migrate all the code into a side-channel free implementation. The second approach is to introduce randomized behaviors during the execution to obfuscate the information that an attacker can retrieve. One example is cryptographic blinding~\cite{coron1999resistance}. Developers can use a random value to blend the secret key but still generate the same encrypted data with a carefully designed algorithm.  Binding values can alter the program execution to some unpredictable states and introduce the randomization to the execution. However, the binding value can also be retrieved by an attacker through other methods, including side-channel attacks. An attacker can launch an attack to retrieve the binding value to recover the secret input.

\subsubsection*{New Implementation}
Recent work~\cite{bernstein2012security,libsodium,zinzindohoue2017hacl,monocypher} proposes to design and implement clean-slate side-channel free libraries from scratch. Researchers~\cite{bernstein2008chacha,langley2016chacha20} design and implement new constant-time cryptographic algorithms (e.g., ChaCha20) to defend against timing side-channel attacks. Unfortunately, writing side-channel resistant cryptography code and validating the security property can be tedious and error-prone.  Computer-aided cryptography~\cite{barbosa2021sok} is proposed to address the challenge, which encompasses formal verification methods to design, analyze and implement the cryptography libraries. Formal verification on side-channel leakages starts with a leakage model, which abstracts what an attacker can observe during an attack. After defining a leakage model, the implementation is secure (under the leakage model) to side-channel attacks if and only if it can be proved that observations are independent of the input secrets. The property is an example of \textit{noninterference} or \textit{secure information flow}. Eldib et al.~\cite{eldib2014smt} propose using a SMT solver based method to formally verifying masked implementations. Bayrak et al.~\cite{bayrak2013sleuth} verify power-based side-channel countermeasures by transferring the verification problem into a set of Boolean satisfiability problems. However, achieving noninterferecne for real-world software is often impractical and researchers also seek to compromised solutions such as quantitative information flow (see Section~\ref{sec:qif}).

\subsection{Hardware Countermeasures}
\subsubsection*{Resource Partitioning}
Side-channel attacks usually require attackers and victims to share some hardware resources (e.g., CPU cache, MMU). A straightforward way to defend these attacks is to partition the resources and ensure the attacker and the victim never share the same hardware resource~\cite{Page2005PartitionedCA,shi2011limiting,liu2016catalyst}. For example, in a cloud computing environment, several virtual machines (VMs) can be hosted on the same physical machine and these VMs share the same last-level cache (LLC). Researchers proposed different methods to partition the LLC such as cache coloring~\cite{shi2011limiting}, using Intel Cache Allocation Technology~\cite{liu2016catalyst}. Recent work also uses machine checkable methods to verify side-channel resistance in hardware~\cite{Zhang:2015:HDL:2775054.2694372,Li:2014:SLH:2541940.2541947,236334}.
\subsubsection*{Randomization}
This type of defense randomizes the behaviors of hardware resources to make it harder for attackers to infer secrets based on side-channel information. For example, Random Permutation Cache ~\cite{wang2007new} uses a dynamically randomized memory to cache mapping table for each different process. Vattikonda et al.~\cite{vattikonda2011eliminating} add noises to the result of the system timer to mitigate timing-based channels.

\subsection{Discussion on Existing Countermeasures}
Despite a lot of hardware countermeasures that have been proposed in academia, few of them have been adopted in real-world production systems. In general, hardware countermeasures usually require new hardware designs and are harder to adopt in practice.  In addition, some hardware methods (e.g., Oblivious RAM) may introduce expensive computation overhead. Therefore, many practical solutions are still software countermeasures. However, software methods are not very general, and it is hard to find many side-channel leakages in large code bases manually. In this dissertation, we propose side-channel analysis methods to address the problem.

\section{Information Quantification}\label{sec:qif}
Proposed by Denning~\cite{robling1982cryptography} and Gray~\cite{gray1992toward},
Quantitative Information Flow (QIF) aims to provide an estimation of the amount of leaked information from the sensitive information given the public output. If zero bits
of the information are leaked, the program is called non-interference. 
McCamant and Ernst~\cite{McCamantE2008} quantify the information leakage as the network
flow capacity. They model the possible information leakage as a network of limited-capacity
channel. After that, they use the maximum rate of the channel to represent the amount of secret information that the execution can reveal.
Backes et al.~\cite{5207642} propose an automated method for QIF
by computing an equivalence relation on the set of input keys. However,
the evaluation shows the approach only works on small programs.
Besides, their approach cannot handle programs with bitwise operations.
However, such bitwise operations are quite common in cryptographic libraries.
Phan et al.~\cite{Phan:2012:SQI:2382756.2382791} propose symbolic QIF. The goal of their
work is to ensure a program is non-interference. They adopt an over-approximation method to estimate the total information leakage, and their method
does not work for secret-dependent memory access side-channels.
Pasareanu et al.~\cite{pasareanu2016multi} combine symbolic analysis and Max-SMT solving to synthesize the concrete public input that can lead to the worst-case leakage. They assume the target program has multiple different input secrets and calculate the average leakage for one-fixed public input.
CHALICE~\cite{Chattopadhyay:2017:QIL:3127041.3127044} quantifies the leaked
information for a given cache behavior.
It symbolically reasons about cache
behavior and estimates the amount of leaked information based on cache miss/hit.
Their approach only scales to small programs, which limits its usage in
real-world applications. On the contrary, \tool{}~\cite{bao2021abacus} assesses the sensitive level
of side channels with different granularities. It can also analyze side channels in real-world cryptographic libraries.

@@

\section{Trace-based Binary Analysis}
Since our work relies on binary analysis, we briefly mention the related work here.
Trace-based binary analysis can be seen as a type of dynamic analysis. It analyzes one concrete program execution path each time rather than trying to cover
as many paths as possible. Compared to static analysis, trace-based analysis
is usually more precise as it can incorporate runtime information. It usually has
two steps. The first step is to collect trace information. Existing
approaches often rely on binary instrumentation frameworks or emulators (e.g., Intel Pin~\cite{luk2005pin},
Valgrind~\cite{nethercote2007valgrind}, and QEMU). The second step to analyze the trace using techniques, such
as symbolic execution and taint analysis. Depending on when the analysis is launched,
there are two kinds of trace-based analysis, online analysis and offline analysis. As the name suggests, online analysis is performed during the program execution. The target program's data can be modified directly in memory, which is useful to tasks like fuzzing because the analysis
can force the program to execute a specific branch. 
Offline analysis or post-analysis happens after the program finishes executions. 
Trace-based binary analysis has been adopted in several tasks like algorithm
detection, vulnerability detection, and malware analysis. We only mention a few of
the most related works here. Xu et al.~\cite{xu2017cryptographic} introduce bitwise symbolic execution, which can
identify crypto algorithms in obfuscated binaries. Ming et al.~\cite{ming2017binsim} use system call sliced segment equivalence checking to do binary diffing. Jia et al.~\cite{jia2017towards}
perform an offline analysis on execution traces to identify heap overflows.

There are several known binary analysis frameworks that can help users to launch analysis on binary executables. BitBlaze~\cite{brumley2011bap} is a binary analysis platform that provides both static and dynamic analysis. It uses the tracecap plugin to collect execution traces and convert the trace into Vine IR. Binary Analysis Platform (BAP)~\cite{10.1007/978-3-642-22110-1_37}
is another binary analysis platform. Similar to BitBlaze, it uses a Pintrace tool to
collect execution traces and convert the trace into BAP IR. Angr~\cite{shoshitaishvili2016state}, the most popular
binary analysis platform recently, also supports trace-based analysis. It converts each instruction into VEX IR and interprets each IR in Python.

The above existing well-known binary analysis frameworks adopt the IR layer, which makes it easy to implement the analysis and support multiple architectures. However, it can have very low efficiency and long traces. All IR must be interpreted no matter if symbolic variables are involved or not. As side-channels are low-level problems in general, the IR will lose some information. For example, IR-level branches are not necessarily a super-set or a sub-set of machine-code level branches. Some IR-level branches can be compiled into non-branch instructions like conditional moves. The above limitations motivate us to implement a new trace-based binary analysis framework.

>>>>

Table~\ref{table:side_channel_detection_tool} covers some representations of existing side-channel leakage detection tools. 
CacheAudit~\cite{182946} uses abstract domains to compute an
over-approximation of cache-based side-channel information leakage upper bound.
However, it is difficult to judge the sensitive level of the side-channel leakage based on the leakage provided by CacheAudit. Besides, CacheAudit is not scalable enough to analyze ciphers like RSA implementations. CacheS~\cite{236338} improves the approach from
CacheAudit with a new abstract model that only tracks
secret-related code. Like CacheAudit, CacheS cannot
measure the amount of leaked information in each side-channel vulnerability.
CaSym~\cite{Brotzman19Casym} introduces a static cache-aware symbolic reasoning technique to cover multiple paths in a target program. CaSym can work on two threat models: trace-based attacks and access-based attacks. Similarly, their
approaches cannot evaluate the sensitive level for each side-channel vulnerability, and it only works on small code snippets.

@@

\subsubsection*{Software Patch}
Once we have identified side-channel leakages, there are two common approaches to mitigate side-channel attacks. The first approach is to adopt unified behavior code when computing the secret data. For example, a control flow leakage happens when different secret inputs lead to different paths. Software developers can mitigate the leakage by executing both branches and select the result with bit-wise operations~\cite{Coppens:2009:PMT:1607723.1608124}. However, for exiting code base, it is hard to migrate all the code into a side-channel free implementation. The second approach is to introduce randomized behaviors during the execution to obfuscate the information that an attacker can retrieve. One example is cryptographic blinding~\cite{coron1999resistance}. Developers can use a random value to blend the secret key but still generate the same encrypted data with a carefully designed algorithm.  Binding values alters the program execution to some unpredictable states and introduce the randomization to the execution. However, the binding value can also be retrieved by an attacker through other methods, including side-channel attacks. An attacker can launch an attack to retrieve the binding value to recover the secret input.

\subsubsection*{New Implementation}
Recent work~\cite{bernstein2012security,libsodium,zinzindohoue2017hacl,monocypher} proposes to design and implement clean-slate side-channel free libraries from scratch. Researchers~\cite{bernstein2008chacha,langley2016chacha20} design and implement new constant-time cryptographic algorithms (e.g., ChaCha20) to defend against timing side-channel attacks. Unfortunately, writing side-channel resistant cryptography code and validating the security property using pen-and-paper arguments can be tedious and error-prone.  Computer-aided cryptography~\cite{barbosa2021sok} is proposed to address the challenge, which encompasses formal verification methods to design, analyze and implement the cryptography libraries. Formal verification on side-channel leakages starts with defining a leakage model, which abstracts what an attacker can observe during an attack. After defining a leakage model, the implementation is secure (under the leakage model) to side-channel attacks if and only if it can be proved that observations are independent of the input secrets. The property is an example of \textit{noninterference} or \textit{secure information flow}. Eldib et al.~\cite{eldib2014smt} propose using a SMT solver based method to formally verifying masked implementations. Bayrak et al.~\cite{bayrak2013sleuth} verify power-based side-channel countermeasures by transferring the verification problem into a set of Boolean satisfiability problems. However, achieving noninterferecne for real-world software is often impractical and researchers also seek to compromised solutions such as quantitative information flow (see Section~\ref{sec:qif}).

\subsection{Hardware Countermeasures}
\subsubsection*{Resource Partitioning}
Side-channel attacks usually require attackers and victims to share some hardware resources (e.g., CPU cache, MMU). A way to defend these attacks is to partition the resources and ensure the attacker and the victim never share the same hardware resource~\cite{Page2005PartitionedCA,shi2011limiting,liu2016catalyst}. For example, in a cloud computing environment, several virtual machines (VMs) can be hosted on the same physical machine and these VMs share the same last-level cache (LLC). Researchers proposed different methods to partition the LLC such as cache coloring~\cite{shi2011limiting}, using Intel Cache Allocation Technology~\cite{liu2016catalyst}. Recent work also uses machine checkable methods to verify side-channel resistance in hardware~\cite{Zhang:2015:HDL:2775054.2694372,Li:2014:SLH:2541940.2541947,236334}.
\subsubsection*{Randomization}
This type of defense randomizes the behaviors of hardware resources to make it harder for attackers to infer secrets based on side-channel information. For example, Random Permutation Cache ~\cite{wang2007new} uses a dynamically randomized memory to cache mapping table for each different process. Vattikonda et al.~\cite{vattikonda2011eliminating} add noises to the result of the system timer to mitigate timing side-channel attacks.

\subsection{Discussion on Existing Countermeasures}
Despite a lot of hardware countermeasures that have been proposed in academia, few of them have been adopted in real-world production systems. In general, hardware countermeasures usually require new hardware designs and are harder to adopt in practice.  In addition, some hardware methods (e.g., Oblivious RAM~\cite{fletchery2014suppressing}) may introduce expensive computation overhead. Therefore, many practical solutions are still software countermeasures. However, software methods are not very general, and it is hard to identify many side-channel leakages in large software manually. In this dissertation, we propose side-channel analysis methods to address the problem.

\section{Information Quantification}\label{sec:qif}
Proposed by Denning~\cite{robling1982cryptography} and Gray~\cite{gray1992toward},
Quantitative Information Flow (QIF) aims to provide an estimation of the amount of leaked information from the sensitive information given the public output. 
McCamant and Ernst~\cite{McCamantE2008} quantify the information leakage as the network
flow capacity. They model possible information leakage as a network of limited-capacity
channel. After that, they use the maximum rate of the channel to represent the amount of secret information that the execution can reveal.
Backes et al.~\cite{5207642} propose an automated method for QIF
by computing an equivalence relation on the set of input keys. However,
the evaluation shows the approach only works on small programs.
Besides, their approach cannot handle programs with bitwise operations.
However, such bitwise operations are quite common in cryptographic libraries.
Phan et al.~\cite{Phan:2012:SQI:2382756.2382791} propose symbolic QIF. The goal of their
work is to ensure a program is \textit{noninterference}. They adopt an over-approximation method to estimate the total information leakage, and their method
does not work for secret-dependent memory access side-channels.
Pasareanu et al.~\cite{pasareanu2016multi} combine symbolic analysis and Max-SMT solving to synthesize the concrete public input that can lead to the worst-case leakage. They assume the target program has multiple different input secrets and calculate the average leakage for one-fixed public input.
CHALICE~\cite{Chattopadhyay:2017:QIL:3127041.3127044} quantifies the leaked
information for a given cache behavior.
It symbolically reasons about cache
behaviors and estimates the amount of leaked information based on cache miss/hit.
Their approach only scales to small programs, which limits its usage in
real-world applications. On the contrary, \tool{}~\cite{bao2021abacus} assesses the sensitive level
of side channels with different granularities. It can also analyze side channels in real-world cryptographic libraries.

@@

\section{Trace-based Binary Analysis}
Since our work relies on binary analysis, we briefly mention the related work here.
Trace-based binary analysis can be seen as a type of dynamic analysis. It analyzes one concrete program execution path each time rather than trying to cover
as many paths as possible. Compared to static analysis, trace-based analysis
is usually more precise as it can incorporate runtime information. It usually has
two steps. The first step is to collect trace information. Existing
approaches often rely on binary instrumentation frameworks or emulators (e.g., Intel Pin~\cite{luk2005pin},
Valgrind~\cite{nethercote2007valgrind}, and QEMU~\cite{bellard2005qemu}). The second step to analyze the trace using techniques such
as symbolic execution and taint analysis. Depending on when the analysis is launched,
there are two kinds of approaches, online analysis and offline analysis. As the name suggests, online analysis is performed during the program execution. The target program's data can be modified directly in memory, which is useful to tasks such as fuzzing because the analysis
can force the program to execute a specific branch. 
Offline analysis or post-analysis happens after the program finishes executions. 
Trace-based binary analysis has been adopted in several tasks such as algorithm
detection, vulnerability detection, and malware analysis. We only mention a few of
the most related works here. Xu et al.~\cite{xu2017cryptographic} introduce bitwise symbolic execution, which 
detects crypto algorithms in obfuscated binaries. Ming et al.~\cite{ming2017binsim} use the method called sliced segment equivalence checking to do binary diffing. Jia et al.~\cite{jia2017towards} perform an offline analysis on execution traces to identify heap overflows.

There are several well-known binary analysis frameworks that can help users to perform analysis on binary executables. BitBlaze~\cite{brumley2011bap} is a binary analysis platform that provides both static and dynamic analysis. It uses the tracecap plugin to collect execution traces and convert the trace into Vine IR. Binary Analysis Platform (BAP)~\cite{10.1007/978-3-642-22110-1_37}
is another binary analysis platform. Similar to BitBlaze, it uses a pintrace tool to
collect execution traces and convert the trace into BAP IRs. Angr~\cite{shoshitaishvili2016state}, the most popular
binary analysis platform recently, also supports trace-based analysis. It converts X86 instructions into VEX IRs and interprets each IR in Python.

All above binary analysis frameworks adopt IR layers, which makes it easy to implement the analysis and support multiple architectures. However, it can have very low efficiency and long traces. All IRs must be interpreted no matter symbolic variables are involved or not. As side-channels are low-level problems in general, the IR will lose some information. For example, IR-level branches are not necessarily a super-set or a sub-set of machine-code level branches. Some IR-level branches can be compiled into non-branch instructions like conditional moves. The above limitations motivate us to implement a new trace-based binary analysis framework.
